---
title: "Analytics"
author: <a href="https://schuyler-smith.github.io/" target="_blank" >Schuyler D. Smith</a>
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r load_data, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=4, cache = TRUE)
library(phylosmith)
data(soil_column)
```

<br>

## common_taxa
<a name="common_taxa"></a>

Used to identify which entries in the taxa_table are shared among treatment-groups. It will return a `vector` of taxa names that are all seen in `n` groups.

<br>
***Usage***
```r
common_taxa(phyloseq_obj, treatment = NULL, subset = NULL, n = 'all')
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object. It must contain sample_data() with information about each sample, and it must contain tax_table() with information about each taxa/gene.
`treatment` | Column name as a `string` or `numeric` in the `sample_data`. This can be a `vector` of multiple columns and they will be combined into a new column.
`subset` | A factor within the `treatment`. This will remove any samples that to not contain this factor. This can be a `vector` of multiple factors to subset on.
`n` | Number of treatment groups that need to share the taxa to be considered a common taxa.

***Examples***
```{r common_taxa}
common_taxa(soil_column, treatment = c('Matrix', 'Treatment'), subset = 'Unamended', n = 'all')[1:35]
```

<br>

***

<br>

## taxa_core
<a name="taxa_core"></a>

Filter taxa in phyloseq-object to only include core taxa

***Usage***
```r
taxa_core(phyloseq_obj, treatment = NULL, subset = NULL, frequency = 0.5, abundance_threshold = 0.01)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object. It must contain sample_data() with information about each sample, and it must contain tax_table() with information about each taxa/gene.
`treatment` | Column name as a `string` or `numeric` in the `sample_data`. This can be a `vector` of multiple columns and they will be combined into a new column.
`subset` | A factor within the `treatment`. This will remove any samples that to not contain this factor. This can be a `vector` of multiple factors to subset on.
`frequency` | The proportion of samples the taxa is found in.
`abundance_threshold` | The minimum relative abundance the taxa is found in for each sample.

***Examples***
The `soil_column` data has 18,441 OTUs listed in its `taxa_table`.
```{r taxa_core}
taxa_core(soil_column, frequency = 0.2, abundance_threshold = 0.01)
```

<br>

***

<br>

## taxa_proportions
<a name="taxa_proportions"></a>

Computes the proportion of a taxa classification. This can be done by treatment, sample, or across the dataset.

***Usage***
```r
taxa_proportions(phyloseq_obj, classification, treatment = NA)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object. It must contain sample_data() with information about each sample, and it must contain tax_table() with information about each taxa/gene.
`classification` | Column name as a `string` or `numeric` in the tax_table for the prportions to be reported on.
`treatment` | Column name as a `string` or `numeric` in the `sample_data`. This can be a `vector` of multiple columns and they will be combined into a new column.

***Examples***
```{r taxa_proportions_treatment}
taxa_proportions(soil_column, 'Phylum', treatment = c('Matrix', 'Treatment'))
```
```{r taxa_proportions_sample}
taxa_proportions(soil_column, 'Phylum', treatment = 'sample')
```
```{r taxa_proportions_whole}
taxa_proportions(soil_column, 'Phylum', treatment = NA)
```

<br>

***

<br>

## unique_taxa
<a name="unique_taxa"></a>

Used to identify which entries in the taxa_table are unique to a specific treatment-group. It will return a `list` of `vector`s of taxa names that are only seen in each group.

<br>
***Usage***
```r
unique_taxa(phyloseq_obj, treatment, subset = NULL)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object. It must contain sample_data() with information about each sample, and it must contain tax_table() with information about each taxa/gene.
`treatment` | Column name as a `string` or `numeric` in the `sample_data`. This can be a `vector` of multiple columns and they will be combined into a new column.
`subset` | A factor within the `treatment`. This will remove any samples that to not contain this factor. This can be a `vector` of multiple factors to subset on.

***Examples***
```{r unique_taxa}
unique_taxa(soil_column, treatment = c('Matrix', 'Treatment'), subset = 'Unamended')
```

<br>

***
