---
title: "Analytics"
author: <a href="https://schuyler-smith.github.io/" target="_blank" >Schuyler D. Smith</a>
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r load_package, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=4, cache = TRUE)
library(phylosmith)
data(soil_column)
```

Examples used in this vignette will use the `GlobalPatterns` dataset from `phyloseq`.

```{r load_data, message=FALSE, warning=FALSE}
library(phyloseq)
data(GlobalPatterns)
```

<br>

## Normalization Methods

<br>

### library_size
<a name="library_size"></a>

Performs a library-size normalization on the `phyloseq-object`

<br>
***Usage***
```r
library_size(phyloseq_obj)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.

***Examples***
```{r library_size, cache = TRUE}
phyloseq::sample_sums(GlobalPatterns)
normalized_obj <- library_size(GlobalPatterns)
phyloseq::sample_sums(normalized_obj)
```

<br>

***

<br>

### relative_abundance
<a name="relative_abundance"></a>

Transforms the the `otu_table` count data to relative abundance. Relative abundance sets the count sums for each sample to 1, and then assigns each taxa an abundance equal to its proportion on the total sum (very low abundance taxa may ).

<br>
***Usage***
```r
relative_abundance(phyloseq_obj)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object that contains `otu_table` count data.

***Examples***

```{r relative_abundance}
phyloseq::sample_sums(relative_abundance(GlobalPatterns, 10))
```

<br>

***

<br>

## Taxa

<br>

### common_taxa
<a name="common_taxa"></a>

Used to identify which entries in the taxa_table are shared among 
treatment-groups. It will return a `vector` of taxa names that are all seen
in `n` groups.

<br>
***Usage***
```r
common_taxa(phyloseq_obj, treatment = NULL, subset = NULL, n = 'all')
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to not contain this factor. This can be a `vector` of multiple factors to subset on.
`n` | Number of treatment groups that need to share the taxa to be considered a common taxa.

***Examples***
```{r common_taxa}
common_taxa(GlobalPatterns, treatment = 'SampleType', 
  subset = 'Tongue', n = 'all')[1:35]
```

<br>

***

### taxa_core
<a name="taxa_core"></a>

Filter taxa in phyloseq-object to only include core taxa

***Usage***
```r
taxa_core(phyloseq_obj, treatment = NULL, subset = NULL, frequency = 0.5, abundance_threshold = 0.01)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to not contain this factor. This can be a `vector` of multiple factors to subset on.
`frequency` | The proportion of samples the taxa is found in.
`abundance_threshold` | The minimum relative abundance the taxa is found in for each sample.

***Examples***
The `soil_column` data has 18,441 OTUs listed in its `taxa_table`.
```{r taxa_core}
taxa_core(GlobalPatterns, frequency = 0.2, abundance_threshold = 0.01)
```

<br>

***

### taxa_proportions
<a name="taxa_proportions"></a>

Computes the proportion of a taxa classification. This can be done by treatment, sample, or across the dataset.

***Usage***
```r
taxa_proportions(phyloseq_obj, classification, treatment = NA)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | `phyloseq_obj` | A phyloseq-class object.
`classification` | Column name as a `string` or `numeric` in the tax_table for the prportions to be reported on.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.

***Examples***
```{r taxa_proportions_treatment}
taxa_proportions(GlobalPatterns, 'Phylum', treatment = "SampleType")
```
```{r taxa_proportions_sample}
taxa_proportions(GlobalPatterns, 'Phylum', treatment = 'Sample')
```
```{r taxa_proportions_whole}
taxa_proportions(GlobalPatterns, 'Phylum', treatment = NULL)
```

<br>

***

<br>

### unique_taxa
<a name="unique_taxa"></a>

Identify which taxa are unique to a specific treatment-group. It will return a `list` of `vector`s of taxa-names that are only seen in each group.

<br>
***Usage***
```r
unique_taxa(phyloseq_obj, treatment, subset = NULL)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to not contain this factor. This can be a `vector` of multiple factors to subset on.

***Examples***
```{r unique_taxa}
uniques <- unique_taxa(GlobalPatterns, treatment = "SampleType")
data.frame(lapply(uniques, "length<-", max(lengths(uniques))))
```

<br>

***

