---
title: "Networks"
author: <a href="https://schuyler-smith.github.io/" target="_blank" >Schuyler D. Smith</a>
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r load_package, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=16, fig.height=8, cache = TRUE)
library(phylosmith)
```

Examples used in this vignette will use the `GlobalPatterns` dataset from 
`phyloseq`.

```{r load_data, message=FALSE, warning=FALSE}
library(phyloseq)
data(GlobalPatterns)
```

<br>

## co_occurrence_network
<a name="co_occurrence_network"></a>

Create a ggplot object of the co-occurrence network.

<br>
***Usage***
```r
co_occurrence_network(phyloseq_obj, treatment = NULL, subset = NULL, co_occurrence_table = NULL, classification = NULL, node_colors = 'default', cluster = FALSE, cluster_colors = 'default', buffer = 0.5)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to
not contain this factor. This can be a `vector` of multiple factors to subset on.
`co_occurrence_table` | Table of the co-occurrence of taxa/genes in the phyloseq_obj, computed using `co_occurrence()`. If no table is given, it will be computed with the phyloseq_obj, using the given treatment and p = 0.05 with a rho cut-off of 0.8.
`classification` | Column name as a `string` or `numeric` in the `tax_table` for the factor to use for node colors.
`node_colors` | Name of a color set from the <a href="https://cran.r-project.org/web/packages/RColorBrewer/RColorBrewer.pdf" target="_blank" >`RColorBrewer`</a> package or a vector palete of R-accepted colors.
`cluster` |  if `TRUE`, will use `igraph`'s `cluster_fast_greedy` method. Alternatively, you may pass a vector of cluster assignments with order corresponding to the order of the taxa_names in the `phyloseq_obj`.
`cluster_colors` | Name of a color set from the <a href="https://cran.r-project.org/web/packages/RColorBrewer/RColorBrewer.pdf" target="_blank" >`RColorBrewer`</a> package or a vector palete of R-accepted colors.
`buffer` | Amount of space beyond the points to extend the cluster (aesthetic argument).

***Examples***
This sample set is too small to get any meaningful co-occurrence.
```{r co_occurrence_network}
filtered_obj <- conglomerate_taxa(GlobalPatterns, "Phylum")
co_occurrence_network(filtered_obj, treatment = "SampleType", 
  classification = 'Phylum')
```

<br>

***

<br>

## network_layout_ps
<a name="network_layout_ps"></a>

Create an layout_igraph object of the co-occurrence from a phyloseq object. This can be input into the co_occurrence_network function, or used for other network creating scripts. The purpose is to be able to create reproducible and comparable graphics.

<br>
***Usage***
```r
network_layout_ps(phyloseq_obj, treatment = NULL, subset = NULL, co_occurrence_table = NULL, algorithm = 'fr')
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to
not contain this factor. This can be a `vector` of multiple factors to subset on.
`co_occurrence_table` | Table of the co-occurrence of taxa/genes in the phyloseq_obj, computed using `co_occurrence()`. If no table is given, it will be computed with the phyloseq_obj, using the given treatment and p = 0.05 with a rho cut-off of 0.8.
`algorithm` | Supported `igraph::layout_` algorithm.

***Examples***
```{r network_layout_ps}
filtered_obj <- conglomerate_taxa(GlobalPatterns, "Phylum")
network_layout_ps(filtered_obj, treatment = "SampleType", algorithm = 'kk')
```

<br>

***

<br>

## network_ps
<a name="network_ps"></a>

Create an layout_igraph object of the co-occurrence from a phyloseq object. This can be input into the co_occurrence_network function, or used for other network creating scripts. The purpose is to be able to create reproducible and comparable graphics.

<br>
***Usage***
```rnetwork_ps(phyloseq_obj, treatment = NULL, subset = NULL, co_occurrence_table = NULL, rho = 0.6)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to
not contain this factor. This can be a `vector` of multiple factors to subset on.
`co_occurrence_table` | Table of the co-occurrence of taxa/genes in the phyloseq_obj, computed using `co_occurrence()`. If no table is given, it will be computed with the phyloseq_obj, using the given treatment and p = 0.05 with a rho cut-off of 0.8.
`rho` | Cutoffs to use to subset the `correlation_table` by correlation values.


***Examples***
```{r network_ps}
filtered_obj <- conglomerate_taxa(GlobalPatterns, "Phylum")
network_ps(filtered_obj, treatment = "SampleType", rho = 0.6)
```

<br>

***

<br>

## variable_correlation_network
<a name="variable_correlation_network"></a>

Create a network ggplot object of the correlation of taxa and sample variables from a phyloseq object. 
This may also be refered to as a bipartite graph. Function from the phylosmith-package.

<br>
***Usage***
```r
variable_correlation_network(phyloseq_obj, variables, classification = NULL,
 treatment = NULL, subset = NULL, correlation_table = NULL, method = 'spearman',
 rho_threshold = c(-0.01, 0.01), p_threshold = 0.05, colors = 'default',
 negative_positive_colors = c('pink1', 'gray22'))
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`co_occurrence_table` | Table of the co-occurrence of taxa/genes in the phyloseq_obj, computed using `co_occurrence()`. If no table is given, it will be computed with the phyloseq_obj, using the given treatment and p = 0.05 with a rho cut-off of 0.8.
`variables` | Numerical factors within the in the sample_data to correlate with the abundance data.
`classification` | Column name as a string or numeric in the tax_table for the factor to conglomerate by.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to
not contain this factor. This can be a `vector` of multiple factors to subset on.
`correlation_table` | Table of the correlation of taxa/variables in the phyloseq_ob}, computed using `variable_correlation`. If no table is given, it will be computed with the phyloseq_obj, using the given treatment and p = 0.05.
`method` | Which correlation method to calculate, "pearson", "spearman".
`rho_threshold` | Cutoffs to use to subset the `correlation_table` by correlation values.
`p_threshold` | Cutoffs to use to subset the `correlation_table` by singnificance values.
`colors` | Name of a color set from the <a href="https://cran.r-project.org/web/packages/RColorBrewer/RColorBrewer.pdf" target="_blank" >`RColorBrewer`</a> package or a vector palete of R-accepted colors.
`negative_positive_colors` | colors to use for the edges to represent negative and positive correlations. Defaul ('pink1', 'gray22')


***Examples***
```{r variable_correlation_network}
filtered_obj <- phyloseq::subset_samples(enterotype, 
  !is.na(enterotype@sam_data$ClinicalStatus))
filtered_obj <- taxa_filter(filtered_obj, frequency = 0.65)
variable_correlation_network(filtered_obj, variables = 'Age',
 treatment = "ClinicalStatus", subset = c("healthy", "elderly", "obese"), 
 classification = 'Genus', method = 'spearman')
```

<br>

***

