---
title: "Data Wrangling"
author: <a href="https://schuyler-smith.github.io/" target="_blank" >Schuyler D. Smith</a>
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r load_package, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=4, cache = TRUE)
library(phylosmith)
data(soil_column)
```

Examples used in this vignette will use the `GlobalPatterns` dataset from `phyloseq`.

```{r load_data, message=FALSE, warning=FALSE}
library(phyloseq)
data(GlobalPatterns)
```

<br>

### conglomerate_samples
<a name="conglomerate_samples"></a>

Merges samples within a `phyloseq-class` object which match on the given criteria (`treatment`). Any sample_data factors that do not match will be set to `NA`. `otu_table` counts will be reassigned as the **mean** of all the samples that are merged together. 

*Use this with caution as replicate samples may be crucial to the experimental design and should be proven statistically to be similar enough to combine for downstream analysis.*

<br>
***Usage***
```r
conglomerate_samples(phyloseq_obj, treatment, subset = NULL)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` |  A `phyloseq-class` object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to not contain this factor. This can be a `vector` of multiple factors to subset on.

***Examples***
```{r conglomerate_samples, warning = FALSE, cache = TRUE}
phyloseq::sample_sums(GlobalPatterns)
conglomerated <- conglomerate_samples(GlobalPatterns, treatment = 'SampleType')
phyloseq::sample_sums(conglomerated)
```

<br>

***

<br>

### conglomerate_taxa
<a name="conglomerate_taxa"></a>

A re-write of the `phyloseq::tax_glom()`. This iteration runs faster with the implementation of `data.table`.

<br>
***Usage***
```r
conglomerate_taxa(phyloseq_obj, classification, hierarchical = TRUE)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`classification` | Column name as a `string` in the `tax_table` for the factor to conglomerate by.
`hierarchical` | Whether the order of factors in the tax_table represent a decreasing hierarchy (`TRUE`) or are independant (`FALSE`). If `FALSE`, will only return the factor given by `classification`.

***Examples***
```{r conglomerate_taxa, warning = FALSE, cache = TRUE}
conglomerate_taxa(GlobalPatterns, classification = 'Phylum', hierarchical = TRUE)
```

<br>

***

<br>

### melt_phyloseq
<a name="melt_phyloseq"></a>

Converts the `otu_table`, `tax_table`, and `sam_data` to a 2-dimensional `data.table`.

<br>
***Usage***
```r
melt_phyloseq(phyloseq_obj)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.

***Examples***
```{r melt_phyloseq}
melt_phyloseq(GlobalPatterns)
```

<br>

***

<br>

### merge_treatments
<a name="merge_treatments"></a>

Combines multiple columns from the sample-data into a single column. Doing this can make it easier to subset and look at the data on multiple factors.

<br>
***Usage***
```r
merge_treatments(phyloseq_obj, ...)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object. It must contain sample_data() with information about each sample.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.

***Examples***
```{r merge_treatments}
merge_treatments(GlobalPatterns, c('Final_Barcode', 'Barcode_truncated_plus_T'))
```

<br>

***

<br>

### set_sample_order
<a name="set_sample_order"></a>

Arranged the phyloseq object so that the samples are listed in a given order, or sorted on metadata. This is most useful for visual inspection of the metadata, and having the samples presented in a correct order in `ggplot2` figures.

***Usage***
```r
set_sample_order(phyloseq_obj, treatment)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.

***Examples***
```{r set_sample_order}
phyloseq::sample_names(GlobalPatterns)
ordered_obj <- set_sample_order(GlobalPatterns, "SampleType")
phyloseq::sample_names(ordered_obj)
```

<br>

***

### set_treatment_levels
<a name="set_treatment_levels"></a>

Set the order of the levels of a factor in the sample-data. Primarily useful for
easy formatting of the order that `ggplot2` will display samples.

***Useful for:***

  * managing order which variables appear in figures

***Usage***
```r
set_treatment_levels(phyloseq_obj, treatment, order)
```

<br>
***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`order` | The order of factors in `treatment` column as a `vector` of `string`s. If assigned "numeric" will set ascending numerical order.

***Examples***
```{r set_treatment_levels}
levels(soil_column@sam_data$Day)
ordered_days <- set_treatment_levels(soil_column, 'Day', 'numeric')
levels(ordered_days@sam_data$Day)
```

<br>

***

### taxa_extract
<a name="taxa_extract"></a>

Create a new phyloseq-object containing defined taxa. Taxa names can be a 
substring or entire taxa name. It will match that `string` in all taxa levels
unless a specific `classification` level is declared.

***Useful for:***

  * looking at specific taxa of interest

***Usage***
```r
taxa_extract(phyloseq_obj, taxa_to_extract, classification = NULL)
```

<br>
***Arguments***

Call | Description
-------------------- | ---------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`taxa_to_extract` | A `string`, or `vector` of taxa of interest.
`classification` | Column name as a `string` in the `tax_table` for the factor 
to conglomerate by.

***Examples***
```{r taxa_extract}
GlobalPatterns
taxa_extract(GlobalPatterns, c("Cyano", "Proteo","Actinobacteria"))
```

<br>

***

<br>

### taxa_filter
<a name="taxa_filter"></a>

This is a robust function that is implemented in nearly every other function of this package. It uses many of the subsetting processes distributed within `phyloseq`, but strives to make them a more user-friendly and combined into a one-stop function. The function works in several steps.

  * Checks to see if `treatments` were specified. If so, it splits the phyloseq into separate objects for each treatment to process.
  * Check to see which taxa are seen in a proportion of samples across each phyloseq object > `frequency` (filtering out taxa seen in few samples) and then merge back to one object
  * If `subset` is declared, remove all `treatment` outside of the `subset`
  * If `drop_samples` is `TRUE` then remove any samples that have 0 taxa observed after filtering (this is a very situational need)

If `frequency` is set to 0 (default), then the function removes any taxa with no abundance in any sample.

***Useful for:***

  * subsetting by sample_data factors
  * removing low-presence taxa
  * removing high-presence taxa

***Usage***
```r
taxa_filter(phyloseq_obj, treatment = NULL, subset = NULL, frequency = 0, below = FALSE, drop_samples = FALSE)
```

<br>

***Arguments***

Call | Description
-------------------- | ------------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`treatment` | Column name as a `string`, or `vector` of, in the `sample_data`.
`subset` | A factor within the `treatment`. This will remove any samples that to not contain this factor. This can be a `vector` of multiple factors to subset on.
`frequency` | The proportion of samples the taxa is found in.
`below` | Does frequency define the minimum (`FALSE`) or maximum (`TRUE`) proportion of samples the taxa is found in.
`drop_samples` | Should the function remove samples that that are empty after removing taxa filtered by frequency (`TRUE`).

***Examples***
The `soil_column` data has 19,216 OTUs listed in its `taxa_table`.
```{r check_data}
GlobalPatterns
```
However, 228 of those taxa are not actually seen in any of the samples.
```{r number_emtpy_taxa}
length(phyloseq::taxa_sums(GlobalPatterns)[phyloseq::taxa_sums(GlobalPatterns) == 0])
```
`taxa_filter` with `frequency = 0` will remove those taxa.
```{r taxa_filter_1}
taxa_filter(GlobalPatterns, frequency = 0)
```
Say that we wanted to only look at taxa that are seen in 80% of the samples.
```{r taxa_filter_2}
taxa_filter(GlobalPatterns, frequency = 0.80)
```
But if we want taxa that are seen in 80% of any 1 teatment group;
```{r taxa_filter_3}
taxa_filter(GlobalPatterns, frequency = 0.80, treatment = 'SampleType')
```
It returns a larger number of taxa, since they need to be seen in less samples overall.

<br>

***

### taxa_prune
<a name="taxa_prune"></a>

Create a new phyloseq-object ommitting the defined taxa. Taxa names can be a 
substring or entire taxa name. It will match that `string` in all taxa levels
unless a specific `classification` level is declared.

***Useful for:***

  * removing specific taxa that are not of interest

***Usage***
```r
taxa_prune(phyloseq_obj, taxa_to_remove, classification = NULL)
```

<br>
***Arguments***

Call | Description
-------------------- | ---------------------------------------------------------
`phyloseq_obj` | A phyloseq-class object.
`taxa_to_remove` | A `string`, or `vector` of taxa to remove.
`classification` | Column name as a `string` in the `tax_table` for the factor 
to conglomerate by.

***Examples***
```{r taxa_prune}
GlobalPatterns
taxa_prune(GlobalPatterns, c("Cyano", "Proteo","Actinobacteria"))
```

<br>

***

